Monitoring misinformation related interventions by Facebook, Twitter and YouTube: methods and illustration.

### 2021-08-30 article/draft.tex
YT flags: check for clarity and add source

YouTube may provide an information panel for videos with topics that are prone to misinformation like COVID-19, moon landing, and climate change. 
Information panels provide ressources concerning a potentially controversial topic from independent third parties or authoritative ressources {\color{red} [LIKE blablabla]}. 
Importantly, YouTube states that these panels exist regardless of the point of view expressed in a given video. {\color{red} [REF??]}
%It should be noted that these panels are not yet available in all languages and countries.

@@

To further study the assignment of information panels to videos, we compiled a list of $170$ YouTube videos, fact-checked by Science Feedback between April $2019$ and March $2021$ and marked as containing false or inaccurate information. 
We collected the information panels, when present, in each video by connecting to a US server and scraping the content of the web page. 
Five types of information panels were found in the list of visited videos regarding climate change, COVID-19 Vaccine, COVID-19, vaccine hesitancy and US elections, as shown in figure \ref{youtube_panel_screenshots}.
{\color{red} [ADD Vaccine hesitancy screenshot + REARRANGE SCREENSHOT IN FREQUENCY ORDER]}

\begin{figure}[h]
	\centering
	%\begin{multicols}{1}
		\includegraphics[scale=0.3]{./img/youtube_panels/yt_1.png} 
	%\end{multicols}
	%\begin{multicols}{1}
		\includegraphics[scale=0.29]{./img/youtube_panels/yt_2.png} 
	%\end{multicols}
	%\begin{multicols}{1}
		\includegraphics[scale=0.294]{./img/youtube_panels/yt_3.png}
	%\end{multicols}
	%\begin{multicols}{1}
		\includegraphics[scale=0.3]{./img/youtube_panels/yt_4.png} 
	%\end{multicols}
	\caption{Examples of YouTube information panels displayed under Youtube videos respectively about: \href{https://www.youtube.com/watch?v=OwqIy8Ikv-c}{Climate change}, \href{https://www.youtube.com/watch?v=9sW0OmzcmL0}{COVID-19 Vaccine}, \href{https://www.youtube.com/watch?v=usyQgPU-VrI}{COVID-19} and \href{https://www.youtube.com/watch?v=8PAwc8nlE\_Q}{US elections}. The videos were accessed on July 21, 2021 from a US server.}
	\label{youtube_panel_screenshots}
\end{figure}

@@

We found that almost half (44 \%) of the videos containing misinformation actually had no information panel (Table \ref{youtube_panel_proportions}).
For the rest of videos, the most frequent information panels found concerned COVID-19 (42 \%) and COVID-19 vaccine (11 \%).
It is possible that YouTube is currently focusing its efforts on the current `hot' controversial topics with a potentially wide public impact, leaving room for misinformation related to other topics to spread, but more research is needed to formally compare misinformation videos about COVID-19 and about other subjects.

@@

In our list, we noticed that the same video was sometimes uploaded multiple times and thus appeared on different YouTube links. 
One YouTube link might contain an information panel while its duplicates did not (see figure \ref{duplicates_yt} for an example).  
In addition, we noticed that some keywords in the video titles were often associated with an information panel, such as `COVID-19', `coronavirus', `pandemic' and `testing', while some notational variants like `C O V I D 19' or `Cv19' were not associated with a panel.
Therefore, we suspect that YouTube is automatically adding information panels based on the presence of certain keywords in the video title (and maybe in the video description), independently of the content of the video.
Note that in figure \ref{duplicates_yt}, the video with `C.O.V.I.D.19' in its title had an information panel, hinting that YouTube might adapt its list of `suspect' keywords as new notational variants for COVID-19 appear.

\begin{figure}[h]
	\centering
		\includegraphics[scale=0.24]{./img/youtube_panels/graph_no_banner.png} 
		\includegraphics[scale=0.24]{./img/youtube_panels/graph_with_banner.png} 
		%\includegraphics[scale=0.2]{./img/youtube_panels/news_with_banner.png}
		%\includegraphics[scale=0.2]{./img/youtube_panels/news_without_banner.png} 
	\caption{Screenshots showing two duplicates of the same video, taken on July 21, 2021. 
Left panel: a \href{https://www.youtube.com/watch?v=d9GbVZOcT18}{YouTube video} about COVID-19 without an information panel. 	
Right panel: a \href{https://www.youtube.com/watch?v=A4RvrEKoNxc}{duplicate of the same video} uploaded under a different title and with a `COVID-19 vaccine' information panel.}
	\label{duplicates_yt}
\end{figure}

>>>>

YouTube provide information panel at the top of search results or under a video related to topics prone to misinformation.
Information panels provide more context on a controversial matter, with a link to an independent, third-party partner's website such as Wikipedia or the World Health Organization. 
Importantly, YouTube states that these panels exist regardless of the point of view expressed in a given video.\footnote{https://support.google.com/youtube/answer/9004474}

@@

To further study the assignment of information panels to videos, we compiled a list of $170$ YouTube videos, fact-checked by Science Feedback between April $2019$ and March $2021$ and marked as containing false or inaccurate information. 
We collected the information panels, when present, under each video by connecting to a US server and scraping the content of the web page. 
Five types of information panels were found regarding climate change, COVID-19 Vaccine, COVID-19, vaccine hesitancy and US elections, as shown in figure \ref{youtube_panel_screenshots}.
{\color{red} [ADD Vaccine hesitancy screenshot]}

\begin{figure}[h]
	\centering
		\includegraphics[scale=0.3]{./img/youtube_panels/yt_1.png} 
		\includegraphics[scale=0.29]{./img/youtube_panels/yt_2.png} 
		\includegraphics[scale=0.294]{./img/youtube_panels/yt_3.png}
		\includegraphics[scale=0.3]{./img/youtube_panels/yt_4.png} 
	\caption{Examples of YouTube information panels displayed under Youtube videos respectively about: \href{https://www.youtube.com/watch?v=OwqIy8Ikv-c}{Climate change}, \href{https://www.youtube.com/watch?v=9sW0OmzcmL0}{COVID-19 Vaccine}, \href{https://www.youtube.com/watch?v=usyQgPU-VrI}{COVID-19} and \href{https://www.youtube.com/watch?v=8PAwc8nlE\_Q}{US elections}. The videos were accessed on July 21, 2021 from a US server.}
	\label{youtube_panel_screenshots}
\end{figure}

@@

Almost half (44 \%) of the videos containing misinformation were shown without any information panel (Table \ref{youtube_panel_proportions}).
For the rest of videos, the most frequent information panels concerned COVID-19 (42 \%) and COVID-19 vaccine (11 \%).
It is possible that YouTube is currently focusing its efforts on the current `hot' controversial topics with a potentially wide public impact, leaving room for misinformation related to other topics to spread, but more research is needed to formally compare misinformation videos about COVID-19 and about other subjects.

@@

We noticed that the same video was sometimes uploaded multiple times and thus appeared on different YouTube links. 
One YouTube link might contain an information panel while its duplicates did not (see figure \ref{duplicates_yt} for an example).  
In addition, we noticed that some keywords in the video titles were often associated with an information panel, such as `COVID-19', `coronavirus', `pandemic' and `testing', while some notational variants like `C O V I D 19' or `Cv19' were not associated with a panel.
Therefore, we suspect that YouTube is automatically adding information panels based on the presence of certain keywords in the video title (and maybe in the video description), independently of the content of the video.
Note that in figure \ref{duplicates_yt}, the video with `C.O.V.I.D.19' in its title had an information panel, hinting that YouTube might adapt its list of suspect keywords as new notational variants for COVID-19 appear, or that some information panels might be added manually (by content regulators or when a video is flagged by many users).

\begin{figure}[h]
	\centering
		\begin{multicols}{1}
		\includegraphics[scale=0.24]{./img/youtube_panels/graph_no_banner.png} &
		\includegraphics[scale=0.24]{./img/youtube_panels/graph_with_banner.png} 
		\end{multicols}
		%\includegraphics[scale=0.2]{./img/youtube_panels/news_with_banner.png}
		%\includegraphics[scale=0.2]{./img/youtube_panels/news_without_banner.png} 
	\caption{Screenshots showing two duplicates of the same video, taken on July 21, 2021. 
Left panel: a \href{https://www.youtube.com/watch?v=d9GbVZOcT18}{YouTube video} about COVID-19 without an information panel. 	
Right panel: a \href{https://www.youtube.com/watch?v=A4RvrEKoNxc}{duplicate of the same video} uploaded under a different title and with a `COVID-19 vaccine' information panel.}
	\label{duplicates_yt}
\end{figure}
